---
title: "Supplementary Materials"
author: "Rob Challen"
date: "24/01/2021"
output: pdf_document
---

```{r setup, include=FALSE}

library(tidyverse)
library(patchwork)
library(survival)
library(survminer)

# optional dependency:

if (dir.exists("~/Git/standard-print-output/")) {
  devtools::load_all("~/Git/standard-print-output/")
  standardPrintOutput::setDefaults()
  spo = TRUE
} else {
  spo = FALSE
}

# devtools::load_all("~/Git/uk-covid-datatools/")
# ukcovidtools::reload()

# load functions
source("./new-variant-mortality.R")

```

```{r}
## Do the analysis ----
if(!exists("coxData3")) 
  coxData3 = loadData(earliestDate = -Inf,minAge = 0)
coxData3 = coxData3 %>% interpretSGene(S_CT=30)

p1 = ggplot(coxData3,aes(x=time,colour=sGene,linetype=deathStatus))+geom_line(stat="count")+scale_y_log10()
p2 = ggplot(coxData3 %>% filter(status==0),aes(x=time,colour=sGene))+ stat_ecdf(geom = "step")+ylab("P(censored before t|censored)")
p3 = ggplot(coxData3 %>% filter(status==1),aes(x=time,colour=sGene))+ stat_ecdf(geom = "step")+ylab("P(died before t|died)")

p1+p2+p3+patchwork::plot_annotation(tag_levels = "A")+patchwork::plot_layout(guides="collect")
```

# Supplementary materials

```{r}

# devtools::load_all("~/Git/uk-covid-datatools/")
# ukcovidtools::reload()
# # grab some ethnicity data. Not going to be used at this point.
# tmp2 = dpc$demog$getDemographicsWithEstimatedEthnicity()
# tmp3 = tmp2 %>% filter(is.finite(count)) %>% inner_join(dpc$codes$getTransitiveClosure() %>% filter(fromCodeType=="LSOA",toCodeType=="NHSER"),by=c("code"="fromCode")) %>% group_by(toCode,ethnicity_final) %>% summarise(count=sum(count))
# popByRegion = tmp3 %>% rename(code = toCode, population = count) #%>% dpc$codes$findNamesByCode()
# popByRegion %>% write_csv("~/Git/vui-202012-1/popDataByEthnicity.csv")
# 
# # 
# # # get overall population by region
# tmp = dpc$demog$getDetailedDemographics()
# popData = tmp %>% semi_join(dpc$codes$getTransitiveClosure() %>% filter(toCode=="E92000001"),by=c("code"="fromCode")) %>% group_by(age) %>% summarise(count = sum(count)) %>% mutate(percent = count/sum(count))
# popData %>% write_csv("~/Git/vui-202012-1/popData.csv")

popByRegion = readr::read_csv("~/Git/vui-202012-1/popDataByEthnicity.csv")
popData = readr::read_csv("~/Git/vui-202012-1/popData.csv")
coxData3 = coxData3 %>% left_join(popByRegion, by=c("NHSER_code"="code","ethnicity_final"),suffix=c("",".ethn"))
```

## Age distributions of cases by sGene status

```{r}



fig1 = ggplot(coxData3,aes(fill=sGeneEra,x=age))+
  geom_bar(aes(y = ..prop..), stat="count",position=position_dodge(width=0.7), width = 0.6)+
  geom_line(data=popData,aes(x=age,y=percent),colour="grey60",inherit.aes = FALSE)+
  ylab("probability")+facet_wrap(~sGeneEra, scales = "free_y", ncol=1)
  

fig1 %>% standardPrintOutput::saveHalfPageFigure("~/Dropbox/covid19/new-variant/ProportionByAgeGroup")
  
```

## Age distributions of infections by sGene status who died

```{r}
fig2 = ggplot(coxData3 %>% filter(diedWithin28days),aes(fill=sGeneEra,x=age))+
  geom_bar(aes(y = ..prop..), stat="count",position=position_dodge(width=0.7), width = 0.6)+
  ylab("probability")+facet_wrap(~sGeneEra, scales = "free_y", ncol=1)
  

fig2 %>% standardPrintOutput::saveThirdPageFigure("~/Dropbox/covid19/new-variant/ProportionDeathsByAgeGroup")
```

## Summary of full unmatched dataset

```{r}

combinedSummary = summaryTable(coxData3) %>% rename_with(~paste0("All ",.x),.cols=c(N,`%age`,`mean (SD)`)) %>%
  left_join(summaryTable(coxData3 %>% filter(sGene != "Unknown")) %>% rename_with(~paste0("Tested ",.x),.cols=c(N,`%age`,`mean (SD)`)),by=c("category","value")) %>%
  left_join(summaryTable(coxData3 %>% filter(diedWithin28days)) %>% rename_with(~paste0("Died ",.x),.cols=c(N,`%age`,`mean (SD)`)),by=c("category","value")) %>%
  left_join(summaryTable(coxData3 %>% filter(diedWithin28days & sGene != "Unknown")) %>% rename_with(~paste0("Died + Tested ",.x),.cols=c(N,`%age`,`mean (SD)`)),by=c("category","value"))

combinedSummary %>% group_by(category) %>% standardPrintOutput::saveTableLandscape("~/Dropbox/covid19/new-variant/CombinedSummaryTable",defaultFontSize = 8)
```

## Cases by IMD

```{r}
s1 = ggplot(coxData3 %>% mutate(
    imd_decile=as.factor(as.numeric(as.character(imd_decile)))
  ),aes(fill=sGeneEra,x=imd_decile))+
  geom_bar(stat="count",position=position_dodge(width=0.7), width = 0.6)+
  ylab("cases")+facet_wrap(vars(sGeneEra), scales = "free")

s1 %>% standardPrintOutput::saveThirdPageFigure("~/Dropbox/covid19/new-variant/SGenePercentByIMD")
```

## Deaths by IMD

```{r}
s2 = ggplot(coxData3 %>% filter(diedWithin28days) %>% mutate(
    imd_decile=as.factor(as.numeric(as.character(imd_decile)))
  ),aes(fill=sGeneEra,x=imd_decile))+
  geom_bar(stat="count",position=position_dodge(width=0.7), width = 0.6)+
  ylab("deaths")+facet_wrap(vars(sGeneEra), scales = "free")

s2 %>% standardPrintOutput::saveThirdPageFigure("~/Dropbox/covid19/new-variant/SGeneDeathsPercentByIMD")
```

## Cases per 1M by Ethnicity

```{r}
s3 = ggplot(
  coxData3 %>% group_by(sGeneEra,ethnicity_final,NHSER_name,population) %>% summarise(count = n()) %>% group_by(sGeneEra,ethnicity_final) %>% summarise(per1M = sum(count)/sum(population)*1000000, pop = sum(population)),
  aes(fill=sGeneEra,x=ethnicity_final,y=per1M))+
  geom_bar(stat="identity",position=position_dodge(width=0.7), width = 0.6)+
  ylab("cases per 1M")+facet_wrap(vars(sGeneEra), scales = "free")

s3 %>% standardPrintOutput::saveThirdPageFigure("~/Dropbox/covid19/new-variant/SGeneCasesPer1MByEthnicity")
```

## Deaths per 1M by Ethnicity

```{r}
s4 = ggplot(
  coxData3 %>% filter(diedWithin28days) %>% group_by(sGeneEra,ethnicity_final,NHSER_name,population) %>% summarise(count = n()) %>% group_by(sGeneEra,ethnicity_final) %>% summarise(per1M = sum(count)/sum(population)*1000000),
  aes(fill=sGeneEra,x=ethnicity_final,y=per1M))+
  geom_bar(stat="identity",position=position_dodge(width=0.7), width = 0.6)+
  ylab("deaths per 1M")+facet_wrap(vars(sGeneEra), scales = "free")

s4 %>% standardPrintOutput::saveThirdPageFigure("~/Dropbox/covid19/new-variant/SGeneDeathsPer1MByEthnicity")
```

## Test reporting delay versus S Gene status


```{r}
s5 = ggplot(coxData3,aes(fill=sGeneEra,x=reportingDelay))+
  geom_bar(aes(y = ..prop..), stat="count",position=position_dodge(width=0.7), width = 0.6)+
  ylab("probability")+coord_cartesian(xlim=c(0,10))
  # geom_text(aes( label = scales::percent(..prop..,accuracy=1),
  #                  y= ..prop.. ), stat= "count", vjust = 0.5, hjust=-0.1, position = position_dodge(width=0.6),size=3,angle=90)

s5 %>% standardPrintOutput::saveThirdPageFigure("~/Dropbox/covid19/new-variant/ReportingDelayBySGeneCases")
```

## Time to admission versus S Gene status


```{r}
s6 = ggplot(coxData3,aes(fill=sGeneEra,x=admissionDelay))+
  geom_bar(aes(y = ..prop..), stat="count",position=position_dodge(width=0.7), width = 0.6)+
  ylab("probability")+
  #geom_density(mapping=aes(colour=sGene),alpha=0.2)+
  #facet_wrap(vars(ethnicity_final),scales="free_y")+
  coord_cartesian(xlim=c(0,15))+facet_wrap(~sGeneEra, scales = "free_y", ncol=1)

s6 %>% standardPrintOutput::saveHalfPageFigure("~/Dropbox/covid19/new-variant/AdmissionDelayBySGeneCases")
```

## Death within 28 days versus CT values / relative copy numbers

```{r}
p1 = ggplot(coxData3,aes(x=sGeneEra,colour=diedWithin28days,y=CT_N))+geom_boxplot()
p2 = ggplot(coxData3,aes(x=sGeneEra,colour=diedWithin28days,y=relativeCopyNumber))+coord_cartesian(ylim=c(0,100))+geom_boxplot()#+facet_wrap(vars(ethnicity_final))
s7 = p1+p2+patchwork::plot_annotation(tag_levels = "A")+patchwork::plot_layout(guides="collect")

s7 %>% standardPrintOutput::saveThirdPageFigure("~/Dropbox/covid19/new-variant/NGeneCTBySGeneAndDeathStatus")
```