---
title: "Increased hazard of mortality in cases compatible with SARS-CoV-2 variant under investigation 202012/1 - a matched cohort study"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: true
knit: (function(inputFile, encoding,...) {
  rmarkdown::render(
    inputFile,
    encoding = encoding,
    output_dir = "~/Dropbox/covid19/new-variant", output_file=paste0('new-variant-mortality-',Sys.Date(),'.pdf'))
  })
# output:
#   word_document :
#     fig_caption: yes
#     fig_width: 7
# knit: (function(inputFile, encoding,...) {
#   rmarkdown::render(
#     inputFile,
#     encoding = encoding,
#     output_dir = "~/Dropbox/covid19/new-variant", output_file=paste0('new-variant-mortality-',Sys.Date(),'.docx'))
#   })
#TODO: https://www.reed.edu/data-at-reed/software/R/markdown_multiple_reports.html
header-includes:
  \usepackage{float}
  \floatplacement{figure}{H}    

fig_width: 7
fig_height: 5
out.width: "100%"
# bibliography: current-rt.bib
# csl: current-rt.csl
# vignette: >
#  %\VignetteIndexEntry{New Variant Mortality}
#  %\VignetteEngine{knitr::rmarkdown}
#  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}

library(tidyverse)
library(patchwork)
library(survival)
library(survminer)

# optional dependency:

if (dir.exists("~/Git/standard-print-output/")) {
  devtools::load_all("~/Git/standard-print-output/")
  standardPrintOutput::setDefaults()
  spo = TRUE
} else {
  spo = FALSE
}

# devtools::load_all("~/Git/uk-covid-datatools/")
# ukcovidtools::reload()

# load functions
source("./new-variant-mortality.R")

```

```{r}
## Do the analysis ----
if(!exists("coxData")) 
  coxData = loadData()

```

## Main analysis

* We choose small tolerances for specimen date & larger for age.
* The default model includes age as a covariate to adjust for mismatches in age. however matching is not biased by age.

```{r}
defaultAgeTol = 5
defaultSpecTol = 1

defaultModel = list(`S gene + age` = Surv(time,status) ~ sGene+age)

models = list(
  `S gene + age` = Surv(time,status) ~ sGene+age,
  `S gene only`=Surv(time,status) ~ sGene, 
  `S gene + N gene CT`=Surv(time,status) ~ sGene+CT_N, 
  `S gene + age + gender` = Surv(time,status) ~ sGene+age+sex,
  `S gene + all covariates` = Surv(time,status) ~ sGene+age+sex+imd_decile+ethnicity_final #+residential_category
)

sGeneOnly = list(
  `S gene only`=Surv(time,status) ~ sGene
)

```

### 1:1 strategy

* So this basically works and gives a significant HR (1.3-3.5) but this is based on 7 deaths in a sample size of 1400
* More 1:1 matches are possible if tolerances are changed

```{r}

run30none1to1 = runAnalysis(
    coxData, sGeneCtThreshold = 30, ctThreshold = 30,
    ageTolerance = defaultAgeTol,specimenDateTolerance=defaultSpecTol,
    modelFormula = models,resolutionStrategy = "none",ratio.neg = 1,ratio.pos = 1)

run30none1to1$table4Data %>% combineBootstraps() %>% prettyPrintSummary()
run30none1to1$table2Data %>% group_by(category) %>% standardPrintOutput::mergeCells()

```

### Ne selection strategy

* So this basically works and gives a significant HR (1.3-3.5) but this is based on 7 deaths in a sample size of 1400
* More 1:1 matches are possible if tolerances are changed

```{r}

run30none = runAnalysis(
    coxData, sGeneCtThreshold = 30, ctThreshold = 30,
    ageTolerance = defaultAgeTol,specimenDateTolerance=defaultSpecTol,
    modelFormula = models,resolutionStrategy = "none")

run30none$table4Data %>% combineBootstraps() %>% prettyPrintSummary()
run30none$table2Data %>% group_by(category) %>% standardPrintOutput::mergeCells()

```

### node sampling strategy

* So this is our headline estimate and strategy
* The node sampling is as far as we can see the least biased strategy as gives even weight to all nodes.
* 2 methods of calculating HR give us the same value of 1.6 (1.2-2.3)
* This is based on 142 deaths
* 50240 pairs are generated (on average per bootstrap)

```{r}

run30nodeSample = runAnalysis(
    coxData, sGeneCtThreshold = 30, ctThreshold = 30,
    ageTolerance = defaultAgeTol,specimenDateTolerance=defaultSpecTol,
    modelFormula = models,resolutionStrategy = "node sample", bootstraps = 10)

run30nodeSample$table4Data  %>% combineBootstraps() %>% prettyPrintSummary()
run30nodeSample$table2Data %>% group_by(category) %>% standardPrintOutput::mergeCells()
run30nodeSample$table5Data  %>% combineBootstraps() %>% prettyPrintSummary()

```

### edge sampling strategy

* The edge sampling strategy is simpler but theoretical risk of bias towards 
* 44233 nodes; 131 deaths.
* gives the same as the node sampling strategy HR 1.6 (1.1-2.3)

```{r}

run30edgeSample = runAnalysis(
    coxData, sGeneCtThreshold = 30, ctThreshold = 30,
    ageTolerance = defaultAgeTol,specimenDateTolerance=defaultSpecTol,
    modelFormula = models,resolutionStrategy = "edge sample", bootstraps = 10)

run30edgeSample$table4Data  %>% combineBootstraps() %>% prettyPrintSummary()
run30edgeSample$table2Data %>% group_by(category) %>% standardPrintOutput::mergeCells()

```

## Central estimates

* Selecting the node sampling strategy as above but with more bootstraps

```{r}

if(!exists("run30")) 
  run30 = runAnalysis(
    coxData, sGeneCtThreshold = 30, ctThreshold = 30,
    ageTolerance = defaultAgeTol,specimenDateTolerance=defaultSpecTol,
    modelFormula = models,resolutionStrategy = "node sample", bootstraps = 100,
    includeRaw=TRUE,includeMatched = TRUE)
  

```

## Pair matching

```{r}

tmp = run30$pairedMatchedUnfilteredData %>% group_by(order.neg,order.pos) %>% summarise(n=n(), meanDate = mean((as.numeric(specimen_date.neg)+as.numeric(specimen_date.pos)/2)))
midpoint = as.numeric((min(tmp$meanDate))+as.numeric(max(tmp$meanDate)))/2
ggplot(tmp,aes(x=order.neg,y=order.pos,fill=n)) + geom_tile()# +geom_label()
ggplot(tmp %>% filter(order.neg<11,order.pos<11), aes(x=order.neg,y=order.pos,fill=n,label=n)) + geom_tile() +geom_text(colour="white")
ggplot(tmp, aes(x=order.neg,y=order.pos,fill=as.numeric(meanDate),label=n)) + geom_tile()+scale_fill_gradient2(low="purple",high="green",mid="yellow",midpoint = midpoint)
```





