---
title: "Increased hazard of mortality in cases compatible with SARS-CoV-2 variant under investigation 202012/1 - a matched cohort study"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: true
knit: (function(inputFile, encoding,...) {
  rmarkdown::render(
    inputFile,
    encoding = encoding,
    output_dir = "~/Dropbox/covid19/new-variant", output_file=paste0('new-variant-mortality-',Sys.Date(),'.pdf'))
  })
header-includes:
  \usepackage{float}
  \floatplacement{figure}{H}    
fig_width: 7
fig_height: 5
out.width: "100%"
---

```{r setup, include=FALSE}

library(tidyverse)
library(patchwork)
library(survival)
library(survminer)

# optional dependency:

if (dir.exists("~/Git/standard-print-output/")) {
  devtools::load_all("~/Git/standard-print-output/")
  standardPrintOutput::setDefaults()
  spo = TRUE
} else {
  spo = FALSE
}

options(dplyr.summarise.inform = FALSE)

# load functions
source("./new-variant-mortality.R")
datadate="20210212"
outdir = paste0("~/Dropbox/covid19/new-variant/",datadate,"/")
if(!dir.exists(outdir)) dir.create(outdir, recursive = TRUE)
outfile = function(x) paste0(outdir,x)
```

```{r}
## Do the analysis ----
if(!exists("coxData")) 
  coxData = loadData(date = datadate, latestDate = as.Date("2021-01-28"))
coxData %>% p_status() %>% p_flowchart(outfile("DataFlow1"))
```

## Main analysis

* We choose small tolerances for specimen date & larger for age.
* The default model includes age as a co-variate to adjust for mismatches in age. however matching is not biased by age.

```{r}
defaultAgeTol = 5
defaultSpecTol = 1

defaultModel = list(
  `S gene + age` = Surv(time,status) ~ sGene+ageDec,
  `S gene + N gene CT + age`=Surv(time,status) ~ sGene+CT_N_10+ageDec
)

models = list(
  `S gene only`=Surv(time,status) ~ sGene, 
  `S gene + age` = Surv(time,status) ~ sGene+ageDec,
  `S gene + N gene CT + age`=Surv(time,status) ~ sGene+CT_N_10+ageDec, 
  `S gene + age + gender` = Surv(time,status) ~ sGene+ageDec+sex,
  `S gene + all covariates` = Surv(time,status) ~ sGene+ageDec+sex+imd_decile+ethnicity_final #+residential_category
)

sGeneOnly = list(
  `S gene only`=Surv(time,status) ~ sGene
)

sGeneAge = list(
  `S gene + age` = Surv(time,status) ~ sGene+ageDec
)

```

## Central estimates

* Run the full analysis with defaults
* and node sampling strategy and with more bootstraps

```{r}

if(!exists("run30")) 
  run30 = runAnalysis(
    coxData, sGeneCtThreshold = 30, ctThreshold = 30,
    ageTolerance = defaultAgeTol,specimenDateTolerance=defaultSpecTol,
    modelFormula = defaultModel,resolutionStrategy = "node sample", bootstraps = 50,
    includeRaw=FALSE,includeMatched = TRUE)

run30$coxMatchedData %>% p_flowchart(outfile("Fig1_DataFlow2"))
```

### Median sample KM curve

* We select one bootstrap sample that produced the median HR for Kaplan-meier visualisation
* If we took all bootstraps confidence intervals would be very narrow.
* This is figure 2

```{r}
tmp = run30$table4Data %>% filter(model=="S gene + age") %>% group_by(boot) %>% summarise(mean.HR = mean(HR)) %>% ungroup() 
middleBoot = which.min(abs(tmp$mean.HR-median(tmp$mean.HR)))
kmfit = survminer::surv_fit(Surv(time,status) ~ sGene, run30$coxMatchedData %>% filter(boot==middleBoot))
ggs = survminer::ggsurvplot(kmfit, conf.int = TRUE)
fig1 = ggs$plot+coord_cartesian(ylim=c(0.995,1))+xlab("Time from first positive test")
if(spo) {
  (fig1+standardPrintOutput::defaultFigureLayout()+standardPrintOutput::narrower())  %>% 
    standardPrintOutput::saveSixthPageFigure(outfile("Fig2_KMCurve"))
} else {
  fig1
}
rm(ggs,kmfit,fig1,tmp)
```
### Summary table

* compares the S neg, Spos and deaths.
* this is table 1

```{r}

if(spo) {
  run30$table2Data %>% group_by(category) %>% 
    standardPrintOutput::saveTable(outfile("Tab1_CaseMatchedSummaryTable"),defaultFontSize = 8)
} else {
  knitr::kable(run30$table2Data)
}
```

### HR table

* this is table 2

```{r}
summary = summariseAnalysisRun(run30) %>% filter(model %in% c("S gene + age", "S gene + N gene CT + age"))
if(spo) {
  prettyPrintSummary(summary) %>% 
    standardPrintOutput::saveTable(outfile("Tab_2HRSummaryTable"),defaultFontSize = 8)
} else {
  knitr::kable(prettyPrintSummary(summary))
}

```

## Biases in matched data set

* any evidence of residual bias?
* this is figure 3

```{r}
## Check for residual biases in case matched data set ----
# small mean age difference and specimen date difference
# no obvious pattern in admission delays
# CT values lower in sGene negative and in those who died. This is ? feature of sGene neg infection
# specimen dates reasonably distributed in time given that the sGene negative numbers increasing. Could constrain analysis to post Dec only.

# There is a mean age difference of `r sprintf("%1.1f",deltas$ageDifference)` years.
# There is a mean specimen date difference of `r sprintf("%1.1f",deltas$specimenDateDifference)` years.

p1 = ggplot(run30$coxMatchedData,aes(fill=sGene,x=admissionDelay))+
  geom_bar(aes(y = ..prop..), stat="count",position=position_dodge(width=0.7), width = 0.6)+
  ylab("probability")+
  xlab("Days to admission")+
  coord_cartesian(xlim=c(0,15))+
  facet_wrap(~sGene, scales = "free_y", ncol=1)+
  scale_fill_brewer(palette="Set1",guide="none")

p2 = ggplot(run30$coxMatchedData,aes(x=sGene,colour=deathStatus,y=CT_N))+geom_boxplot()+ylab("N Gene CT value")+
  scale_color_brewer(palette="Set2")+
  xlab("S gene status")+theme(legend.position = "bottom")

p3 = ggplot(run30$coxMatchedData,aes(fill=sGene,x=specimen_date))+
  geom_bar(stat="count",position=position_dodge(width=0.7), width = 0.6)+
  ylab("count")+
  xlab("Date")+
  facet_wrap(~sGene, scales = "free_y", ncol=1)+scale_fill_brewer(palette="Set1",guide="none")+xlab("date")

fig2 = p1+p3+p2+patchwork::plot_annotation(tag_levels = "A")+patchwork::plot_layout(ncol=3)

if(spo) {
  (fig2+standardPrintOutput::defaultFigureLayout())  %>% 
    standardPrintOutput::saveThirdPageFigure(outfile("Fig3_ResidualBiases"))
} else {
  fig2
}
```

## Sensitivity analysis

* rerun the full analysis for a range of different starting values
* CT values from 20-25
* N.b. this can take some time.

```{r}

if (!exists("summaryCT")) {
  summaryCT = NULL
  for(ct in seq(20,35,1)) {
    run = runAnalysis(coxData, ctThreshold = ct,bootstraps = 10,max=Inf,ageTolerance = defaultAgeTol,specimenDateTolerance=defaultSpecTol,
    modelFormula = sGeneAge,resolutionStrategy = "node sample")
    summaryCT = summaryCT %>% bind_rows(summariseAnalysisRun(run))
  }
}
p1 = ggplot(summaryCT %>% filter(model=="S gene + age" & term == "sGeneNegative"), aes(x=ctThreshold, y=mean.HR,ymin=mean.HR.lower,ymax=mean.HR.upper))+geom_point(colour="red")+geom_errorbar(colour="red")+xlab("CT threshold")+ylab("Hazard ratio")+
  #geom_smooth(method = "lm",formula = y ~ poly(x, 2),se=TRUE,alpha=0.5)
  #geom_smooth(method = "lm",se=TRUE,alpha=0.5)+
  coord_cartesian(ylim=c(0.5,3))+
  gghighlight::gghighlight(ctThreshold==30,unhighlighted_params = list(colour = "black"))


```

* specimen date tolerances from 0 to 14 years

```{r}
if (!exists("summaryDateTol")) {
  summaryDateTol = NULL
  for(dateTol in c(0:14)) {
    run = runAnalysis(coxData, specimenDateTolerance = dateTol,bootstraps = 10,ageTolerance = defaultAgeTol,
    modelFormula = sGeneAge,resolutionStrategy = "node sample")
    summaryDateTol = summaryDateTol %>% bind_rows(summariseAnalysisRun(run))
  }
}

p2 = ggplot(summaryDateTol %>% filter(model=="S gene + age" & term == "sGeneNegative"), aes(x=specimenDateTolerance, y=mean.HR,ymin=mean.HR.lower,ymax=mean.HR.upper))+geom_point(colour="red")+geom_errorbar(colour="red")+xlab("Mismatch tolerance - sample date (days)")+ylab("Hazard ratio")+
  #geom_smooth(method = "lm",formula = y ~ poly(x, 2),se=TRUE,alpha=0.5)
  #geom_smooth(method = "lm",se=TRUE,alpha=0.5)+
  scale_x_continuous(breaks = c(0,7,14))+
  coord_cartesian(ylim=c(0.5,3))+
  gghighlight::gghighlight(specimenDateTolerance==defaultSpecTol,unhighlighted_params = list(colour = "black"))

p4 = ggplot(summaryDateTol %>% filter(model=="S gene + age" & term == "sGeneNegative"),aes(x=specimenDateTolerance,y=mean.specimenDateDifference))+geom_bar(stat="identity", width=0.7, fill="steelblue") + 
  xlab("Mismatch tolerance - sample date (days)")+ylab("Pairwise bias (days)")+scale_x_continuous(breaks = c(0,7,14))

```

* age tolerances from 0 to 9 years

```{r}
if (!exists("summaryAgeTol")) {
  summaryAgeTol = NULL
  for(ageTol in c(0,1,2,3,4,5,6,7,8,9)) {
    run = runAnalysis(coxData, ageTolerance = ageTol,bootstraps = 10,specimenDateTolerance=defaultSpecTol,
    modelFormula = sGeneAge,resolutionStrategy = "node sample")
    summaryAgeTol = summaryAgeTol %>% bind_rows(summariseAnalysisRun(run))
  }
}

p3 = ggplot(summaryAgeTol %>% filter(model=="S gene + age" & term == "sGeneNegative"), aes(x=ageTolerance, y=mean.HR,ymin=mean.HR.lower,ymax=mean.HR.upper))+geom_point(colour="red")+geom_errorbar(colour="red")+xlab("Mismatch tolerance - age (years)")+ylab("Hazard ratio")+
  #geom_smooth(method = "lm",formula = y ~ poly(x, 2),se=TRUE,alpha=0.5)
  #geom_smooth(method = "lm",se=TRUE,alpha=0.5)+
  scale_x_continuous(breaks = c(0,3,6,9))+
  coord_cartesian(ylim=c(0.5,3))+
  gghighlight::gghighlight(ageTolerance==defaultAgeTol,unhighlighted_params = list(colour = "black"))
```

* assemble output of sensitivity analysis
* this is figure 4

```{r}
fig3 = p1+p3+p2+p4+patchwork::plot_annotation(tag_levels = "A")+patchwork::plot_layout(ncol=2)
if(spo) {
  fig3 %>% 
    standardPrintOutput::saveHalfPageFigure(outfile("Fig4_SensitivityAnalysis"))
} else {
  fig3
}
```

